Busch, Khakhalin. Intrinsic temporal tuning

### 2019-05-03 reviews and responses.md

Here you can see peer-reviews for our paper, given in reverse chronological order, and our point-by-point responses to these reviews. Reviews are split into paragraphs, with  responses and comments given after each paragraph.

--------------------------

# Reviews from J Neurophysiol, 2019 Mar 04 (revise and resubmit)

We address the concerns of Reviewer 3 first, as they were overall more critical. Then we respond to the suggestions of Reviewer 2.

@@

We now made the introduction clearer, and better organized, going from the big problem, to specific questions. We moved the presentation of the questions for the study to the very beginning of the second paragraph, and provided a rationale for these questions in the first paragraph.

> Results and Discussion:

> - The rationale for the use of dynamic clamp and the parameters chosen are not clear.

Previously a detailed justification for the use of dynamic clamp was only given in the Discussion section; we now added it to the end of Introduction.

@@

> - I think the experiment in this study is pretty straightforward, but the description of the
results is not. As far as I can tell, tadpoles were exposed to different stimuli (flash,
looming, multisensory, etc.). Then, tectal neurons were stimulated using either 1)
dynamic clamp or 2) more physiologically through optic chiasm. This allowed the
investigators to assess firing when they changed the stimulus intensity and stimulus
duration of the dynamic clamp or the synaptic response to presynaptic stimulation.
They then looked how the different types of sensory experience influenced firing
responses. The rest is data analysis that looks at how the duration, amplitude, total
spiking, and synaptic input is changed in different sensory contexts and sensory
stimuli might interact. I think the text can be made much simpler so the basic design
is easier to grasp. I suggest thinking about what the take home point is for each
figure and craft the results section around those messages.

Thank you; we tried to follow this advice in this revised version of the manuscript!

> - There is also an inconsistent use terminology. For example, “spikiness” and
“excitability” are sometimes used in attempt to describe distinct cell characteristics,
but sometimes it seems like they are being used interchangeably. This makes
interpreting the written results difficult to follow.

In this updated version, we consistently use "spikiness" as a name for the variable, for the first time introduced in Figure 2C. For earlier results section, we replaced "spikiness" with "average number of spikes", and in all instances when we talked about the general idea of neuronal excitability, we called it neuronal excitability. As the only exception to this rule, we opted for the word "spikiness" in the first paragraph of Introduction, to prime the reader for the use of this word later in the text.

@@

> - “Compared to control, neurons from animals exposed to visual stimuli had a flatter
amplitude tuning curve (Fig 1D,E).” What is the reader supposed to be looking at in
the figure? Is this referring to larger conductances for a fixed duration? I think this is
what the authors’ mean, but more intuitive descriptions that matches the figures or
vice versa is needed to make the whole results section easier to digest.

We improved the sentence:

Compared to control, neurons from animals exposed to visual stimuli had a flatter amplitude tuning curve (**less of a difference between columns of Figure 1D, or between curves in each individual panel of Figure 1E**), as they did not increase their spiking as fast in response to larger conductances

@@

> - Excessive wordiness makes the results difficult to read and interpret. For example,
on pg. 9, 4 paragraphs (411 words) are used to describe two graphs, each showing
a different linear relationship (Fig. 3E-F). Why not simply say, “A positive relationship
between temporal tuning and synaptic duration implies a co-tuning of preferr
parameters.”

Of these four paragraphs, the last one reports a negative result that is not shown in Figure 3. The first three paragraphs are indeed dedicated to panels Figure 3EF, but we cannot in good faith replace them with one sentence, as the statement we are making is not statistically obvious. We believe that an attempt to gloss it over by pushing the statistical analysis to the methods section would not be fair. The issue, as we describe in the text, is that while graphs 3E and 3F have similar axes, and both show a linear dependency, they have opposite slopes. Had the slopes be the same, we would have just reported a correlation in one sentence. But with the data we have, we need to be more specific, as we observed that the within-group correlation is positive, yet the between-group correlation is negative: a situation that is typical for larger datasets, but that some readers may not be familiar with.

We are aware that this section is read differently by different readers, and we received some conflicting advice in the past about the level of details needed to communicate this message. For example, those who have explicitly discussed Simpson's paradox in their statistics classes typically find our use of this term very helpful, while those who haven't used this term before, may find it distracting. For now, we opt to include it, even though it seems to be appreciated by only one half of our target audience.

@@

> - In reference to the text at the bottom of page10: Is it possible to include a summary
figure of data from the stepwise generalized linear model in Fig. 4? The purpose of
this part of the study was to try and identify electrophysiological mechanism(s) that
explain the tuning variables. The heart of Fig. 4 should be that they don’t, and figure
that captures this would be helpful.

TODO: How to make a figure like that? They refer to a text about the stepwise generalized linear model. Is there a good visualization here?

> - If these data are interpreted as homeostatic (neurons became less excitable
following sensory experience), is there any evidence that tectal neurons have
different firing activities during the different kinds of sensory experience? For
example, does sound engage tectal neurons less than visual stimuli, and might this
explain some of the trends in the data?

> TODO: that is what we are suggesting in one of the paragraphs. Cite that.

@@

Still, to make this sentence better, we removed the quotes around "preferred", to make it more obvious that it is just our normal word use (we use this word 6 times in the paper), and not an allegory.

> Bottom of page 5, “Effects of acoustic and multisensory stimulation:

> - Is the Auditory stim conditioning referred to as “click sounds” the same as the “startle-inducing
sounds” that is mentioned in the very next sentence? If these are indeed the same stimulus it should
be referred to by the same name, consistently. If they are not the same stimuli, the differences need to
be explained.

These are the same stimuli, and we changes these two sentences to better reflect that:

"To test this question, we exposed tadpoles to four hours of behaviorally salient, startle-inducing “click” sounds, provided at the same frequency (every second) as visual stimuli in the first set of experiments. We found (Fig 1D, E) that exposure to these sounds..."

@@

> - Tadpoles are exposed to 4 hours of click sounds and then intrinsic tuning was measured. No difference
in spiking was observed. This was interpreted to mean that acoustic stimuli did “not strongly activate
tectal circuits during conditioning”. How can the authors know this? Instead, why not determine this
directly by recording click-evoked responses from tectal neurons? Or, if tectal cell responses to auditory
stimuli have been described in previous work, this should be included here.

Unfortunately, recording of tectal reponses to sounds in tadpoles does not seem to be possible in present, not just in our lab (where we don't posess the technical capability to perform in-vivo experiments), but in most labs. Technically, the problem amounts to two limitations: one, to record from a tadpole, one has to paralyze it with a nicotinic receptor anatagonist, which seems to disrupt hearing. Second, even if this limitation is overcome, the sounds has to be strong enough to stimulate the inner ear, but weak enough not to disrupt a recording. Conceivably, both challenges may be overcocme with the use of high speed calcum imaging in freely moving tadpoles, but this is also challenging, and to our knowledge no lab has attempted these recordings yet.

To show that our statement is just a conjecture, we replaced "suggets" with "may suggest" in this sentence: "This may suggest that acoustic stimuli did not activate tectal circuits strongly enough during conditioning..."

@@

> Page 6 line 24: Remove the phrase “are not easily interpretable”. Instead, please interpret your data
clearly.

We now removed this phrase from the main text.

> Page 9: The authors report that the synaptic data show that “prolonged sensory activation with short,
frequent stimuli reshaped synaptic transmission in the tectum, making it faster through selective
potentiation of visual inputs from the eye”.
How do the authors know the effect is selective for RGC inputs? It appears that they only measured
RGC-evoked responses. Auditory/mechanosensory -evoked responses were not reported.

Stimulation in the optic chiam evokes two types of synaptic responses in tectal cells: direct short-latency responses, and delayed recurrent inputs from other cells in the tectum. By looking at average currents early and later after the chiasm shock, we can independently assess the contribution of direct and recurrent inputs. In the paper, in a paragraph before the sentence in question, we report that after stimulation, direct current increased in amplitude, while recurrent currents remained unchanged. This is what we call selective potentiation of direct inputs, as opposed to recurrent inputs. We do not claim that mechanosensory inputs did not change, as we have not recorded auditory responses in this study (recordings of auditory responses are considerably more difficult; see Felch 2016).

To make sure that we are not misunderstood by the reader, we replaced the original sentence with the following two sentences: "This shows that prolonged sensory activation with short, frequent stimuli reshaped synaptic transmission in the tectum, making it faster through selective potentiation of visual inputs from the eye, compared to recurrent inputs within the tectum that remained unchanged. In this study, we did not perform a similar analysis for acoustic projections from the hindbrain."

@@

> Page 9 line 9: “cells exposed to stronger stimuli preferred shorter synaptic inputs in dynamic clamp…”
Here, (where “preferred” is not in quotes) does “preferred” mean to fire more action potentials?

Yes, as we answered above, we now don't use quotation marks for any of the words that aid writing and understanding, so "preferred" consistently means "responded preferably", or "produced stronger responses".

@@

> Page 10 (last paragraph): Were action potential thresholds measured? Were there any differences in
this parameter across groups?

We did not measure action potential thresholds, as we knew from the previous study (Ciarleglio Khakhalin 2015) that potential threshold measurements in tadpole tectal neurons are unreliable, in the sence that they strongly depend on the current that is used to evoke the spikes; this is almost certainly due to rapid Na channels inactivation.

@@

> Page 10 line 8: The “correction for position” is difficult to understand here and in Methods. Why was a
linear regression used – and how is it determined to generate an accurate “correction”? How do the
data look without this correction?

The linear regression is used because it is the simplest way to compensate for a confounding variable (see ANCOVA, or linear models with confounding variables in general). Linear models are the "default" approach in statistics, and are typically used when the exact relation between two variables is unknown, or when the amount of noise is so high that practically there is no difference between linear and conceivable curvilinear regressions. Both situations apply here.

There is no way for us to determine whether our adjustment is "accurate", as there is no experimental way to make more caudal (and thus younger) cells in the tectum older. This situation is not unique however, as it arises in any study that uses linear models, in any field, from ecology to social sciences. When running a study in a human population, if a variable of interest depends on age, one would typically adjust for age by including age as a covariate in the linear model. Here we perform the same standard procedure. After rereading the manuscript, we realized that we used terms "adjustment", "correction", and, in one case, "compensation" interchangeably, which might have led to some confusion. To stress that we are talking about a basic statistical technique of covarate elimination, we now consistently use the term "adjustment" throughout the manuscript.

In the paper, we do not report how the data would have looked without this adjustment for covariates, as it would not be statistically sound to run a linear model with a known covariate without including it. It would underestimate effects for analyses across groups (as the effect of the covarate would be included in the error), and overestimate effects for explanatory linear models in the second half of the paper (as the covariate would contribute to both the dependent, and the independent variables in the model). If the reviewers would like us to numerically re-run any part of the analysis without a position adjustment, please let us know which part you would like to see. Also, all of our technical visualizations and analyses related to this adjustment are available for inspection at: 
https://github.com/khakhalin/Dynamic-clamp-2018/blob/master/dynamic_other_vars.R

### 2019-05-04 reviews and responses.md

Here you can see peer-reviews for our paper, given in reverse chronological order, and our point-by-point responses to these reviews. Reviews are split into paragraphs, with  responses and comments given after each paragraph.

--------------------------

# Reviews from J Neurophysiol, 2019 Mar 04 (revise and resubmit)

The reviews included Reviewer 2, and Reviwer 3, but no Reviewer 1.

@@

These are the same stimuli, and we changes these two sentences to better reflect that:

"To test this question, we exposed tadpoles to four hours of behaviorally salient, startle-inducing “click” sounds, provided at the same frequency (every second) as visual stimuli in the first set of experiments. We found (Fig 1D, E) that exposure to these sounds..."

> - Tadpoles are exposed to 4 hours of click sounds and then intrinsic tuning was measured. No difference
in spiking was observed. This was interpreted to mean that acoustic stimuli did “not strongly activate
tectal circuits during conditioning”. How can the authors know this? Instead, why not determine this
directly by recording click-evoked responses from tectal neurons? Or, if tectal cell responses to auditory
stimuli have been described in previous work, this should be included here.

Unfortunately, recording of tectal reponses to sounds in tadpoles does not seem to be possible at present, not just in our lab (where we don't posess the technical capability to perform in-vivo experiments), but in most labs. Technically, the problem amounts to two limitations: one, to record from a tadpole, one has to paralyze it with a nicotinic receptor anatagonist, which seems to disrupt hearing. Second, even if this limitation is overcome, the sounds has to be strong enough to stimulate the inner ear, but weak enough not to disrupt a recording. Conceivably, both challenges may be overcocme with the use of high speed calcum imaging in freely moving tadpoles, but this is also challenging, and to our knowledge no lab has attempted these recordings yet.

@@

> Page 10 line 8: The “correction for position” is difficult to understand here and in Methods. Why was a
linear regression used – and how is it determined to generate an accurate “correction”? How do the
data look without this correction?

The linear regression is used because it is the simplest way to compensate for a confounding variable (see ANCOVA, or linear models with confounding variables in general). Linear models are the "default" approach in statistics, and are typically used when the exact relation between two variables is unknown, or when the amount of noise is so high that practically there is no difference between linear and conceivable curvilinear regressions. Both situations apply here.

There is no way for us to determine whether our adjustment is "accurate", as there is no experimental way to make more caudal (and thus younger) cells in the tectum older. This situation is not unique however, as it arises in any study that uses linear models, in any field, from ecology to social sciences. When running a study in a human population, if a variable of interest depends on age, one would typically adjust for age by including age as a covariate in the linear model. Here we perform the same standard procedure. After rereading the manuscript, we realized that we used terms "adjustment", "correction", and, in one case, "compensation" interchangeably, which might have led to some confusion. To stress that we are talking about a basic statistical technique of covarate elimination, we now consistently use the term "adjustment" throughout the manuscript.

In the paper, we do not report how the data would have looked without this adjustment for covariates, as it would not be statistically sound to run a linear model with a known covariate without including it. In most cases, it may be expected to underestimate effects for analyses across groups (as the effect of the covarate would be included in the error), and overestimate effects for explanatory linear models in the second half of the paper (as the covariate would contribute to both the dependent, and the independent variables in the model). For each particular comparison, however, both d, F and p values would differ by a small random value. For example, when the measure of "spikness" (one of the three tuning parameters) is compared across groups, a correct analysis with adjustment for position (smean~rostral+medial+group) produces a p-value of 0.01. Without adjustment (smean~group) we get a p-value of 0.03. For temporal tuning, correct analysis (sbend~rostral+medial+group) gives p=0.0304, and if covariate is left unadjusted (sbend~gropu) we get p=0.029. Similarly, for amplitude tuning, we get p=0.00209 for correct analysis, and p=0.0024 for unadjusted analysis. For the analyses in which temporal tuning and ampliltude coefficients were explained through intrinsic parameters, the optimal models were not qualitatively different between adjusted and unadjusted data, but all values in the table were slightly different.

If you would like us to numerically re-run any part of the analysis without a position adjustment, please let us know which part you would like to see. Also, all of our technical visualizations and analyses related to this adjustment are available for inspection at: 
https://github.com/khakhalin/Dynamic-clamp-2018/blob/master/dynamic_other_vars.R

@@

> Page 13, line 7: What does “intrinsic spikiness” mean? Is this another term for intrinsic excitability, or is
this a different parameter? In fact, throughout the manuscript, there are many terms including “intrinsic
excitability, spikiness, number of spikes, and intrinsic spikiness that probably mean the same thing – or
maybe not. If these terms all mean the same thing, then one term should be chosen and used
throughout. If they all refer to different entities, then each one should be clearly defined.

In this updated version, we consistently use "spikiness" as a name for the variable that is for the first time introduced in Figure 2C. For earlier results section, we replaced "spikiness" with "average number of spikes", and in all instances when we talked about the general idea of neuronal excitability, we called it neuronal excitability. As the only exception to this rule, we opted for the word "spikiness" in the first paragraph of Introduction, to prime the reader for the use of this word later in the text.

"Intrinsic spikiness" is no longer used.

"Intrinsic excitability" is used as a name for the phenomenon, that in this particular study we assess mostly using the "spikness" variable, although arguably the number of spikes observed in current clamp experiments can also be considered a different quantification of the same variable. In the text, we say "intrinsic excitability" when we make inferences about the actual properties of the neuron, and "spikiness" when we are talking about the actual value we calculated to assess and quantify this intrinsic excitability.

@@

> Page 15 line 34: The average membrane resistance (330 MOhm) seems a bit low for these small immature
tectal neurons. Were the neurons healthy? What was the average resting membrane potential? Were
there any set criteria for Rm, Vm, or access resistance?

Thank you for noticing that: we have an order of magnitude mistake here (and we are duly ashamed of it). The average membrane resistance was 3.30 GOhm.

TODO: is it even true? Or are these the Gigaseal values, pre-patch?

Smallest membrane resistance observed was 1 GOhm, maximal access resistance 110 MOhm, with the average of 30 MOhm. We did not measure Vm, but measured the holding current necessary to bring the cell to -60 mV; the mean value of this I_hold is -8.16 pA, the resting membrane potential can be estimated from this value with a formula Vm = -60 mV + Rm*I_hold.

All data for all cells is available at:
https://github.com/khakhalin/Dynamic-clamp-2018/blob/master/Data/data_mainInput.txt
(In this dataset, groups "Naive", "SlowF", and "SlowC" should be ignored, as they were not included in the study. Note also that github doesn't show columns correctly, so it may be better to download the file if you want to study it closerly)

@@

> Figure 1 D and E: The average number of spikes fired as a function of conductance duration and all at 3
different peak conductances. Are there any trends at the single cell level? For example, if a given
control neuron fires relatively few spikes in response to 500 ms duration, does it also tend to fire
relatively fewer spikes with 1000ms duration? Or maybe if a given neuron fires a relatively high number
of spikes with 500ms then it will typically fire less at 1000ms duration? Although the individual
responses are shown in 1D, it is impossible to follow the number of spikes fired by any individual neuron
in response to the different conductance durations.

> Figure 4 D, E and F: Is “spikiness” synonymous with “N Spikes”? If these terms do not mean the same
thing, they should be defined. If they do mean the same thing, then one term should be chosen and
used consistently throughout.

Previously a detailed justification for the use of dynamic clamp was only given in the Discussion section; we now moved it to the end of Introduction.

As described in our responses to Reviewer 2, in this updated version, we consistently use "spikiness" as a name for the variable, for the first time introduced in Figure 2C. For earlier results section, we replaced "spikiness" with "average number of spikes", and in all instances when we talked about the general idea of neuronal excitability, we called it neuronal excitability. As the only exception to this rule, we opted for the word "spikiness" in the first paragraph of Introduction, to prime the reader for the use of this word later in the text.

We are not sure how this fact can be illustrated. We searched for papers that use stepwise linear models, and it seems they do not usually include summary figures to describe these models. (They often include tables, listing R-squared, F, df, and p-values for every component, but as in this paper we are not particularly concerned by the precise composition of our models, we doubt a table like that would be useful here). Do you have any specific type of illustration in mind, in which case, could you provide a reference, or maybe describe it? The only type of illustration we could come up with is a scatterplot that compares values predicted from the model to actual values, but it is already shown in the paper, as Figure 4E.

### 2019-05-31 reviews and responses.md

"Intrinsic spikiness" is no longer used.

"Intrinsic excitability" is used as a name for the phenomenon, that in this particular study we assess mostly using the "spikness" variable, although arguably the number of spikes observed in current clamp experiments can also be considered a different quantification of the same variable. In the text, we say "intrinsic excitability" when we make inferences about the actual properties of the neuron, and "spikiness" when we are talking about the actual value we calculated to assess and quantify this intrinsic excitability.

> Page 15 line 34: The average membrane resistance (330 MOhm) seems a bit low for these small immature
tectal neurons. Were the neurons healthy? What was the average resting membrane potential? Were
there any set criteria for Rm, Vm, or access resistance?

Thank you for catching this error! We had an order of magnitude typo here, and the average membrane resistance was in fact 3.2±1.5 GOhm. This value is a bit higher than reported earlier (1.2±0.8 GOhm in Ciarleglio Khakhalin 2015; ~1.1 in Khakhalin 2011), but within what can be reasonably expected for healthy neurons of this size. To make sure that this number reflects our observations, for the purpose of this response, we verified this value by reanalyzing our data offline, and obtained a similar estimation for the average membrane resistance (2.9±1.4 GOhm). We have several reasons to believe that the cells included in this study were healthy. First, the access resistances obtained were good (30±14 MOhm, which is similar to benchmark 50±23 in Ciarleglio Khakhalin 2015, and ~20-30 MOhm in Khakhalin 2011), indicating that higher membrane resistance is not likely to be due to poor access. Second, our neurons reliably produced spikes in current clamp mode, and active currents in response to membrane depolarization in voltage clamp mode, and the quantification of these active conductances yielded numbers similar to those reported before (Ciarleglio Khakhalin 2015). The quality of voltage clamp, as judged by the absence of escaping active currents in IV protocols, was good in the majority of neurons (82% of all neurons recorded). Finally, a few cells that had their membrane resistance change drastically during the recording session, were discarded at the data collection stage. We did not measure the resting potential directly, but measured the holding current necessary to bring the cell to -60 mV in voltage clamp mode, and the mean value of this I_hold value was -8.2±4.8 pA, which is not higher tha what was previously recorded in the literature (19.8±16.6 in Ciearleglio Khakhalin 2015).

Overall, these arguments make us believe that the quality of our patches was good enough to draw the conclusions reported in the study. Note that all data for all cells is available at the following address:
https://github.com/khakhalin/Dynamic-clamp-2018/blob/master/Data/data_mainInput.txt
In this dataset, groups "Naive", "SlowF", and "SlowC" should be ignored, as they were not included in the study. Note also that github doesn't show columns correctly, so if you want to inspect the data, we recommend that you download it and open it in a spreadsheet.

As the mistake your noticed (thank you again for being so thorough!) was a typo in the text, and not a mistake in the data, it does not affect any other calculations reported in this paper.

### 2019-05-31 reviews and responses.md

Overall, these arguments make us believe that the quality of our patches was good enough to draw the conclusions reported in the study. Note that all data for all cells is available at the following address:
https://github.com/khakhalin/Dynamic-clamp-2018/blob/master/Data/data_mainInput.txt
In this dataset, groups "Naive", "SlowF", and "SlowC" should be ignored, as they were not included in the study. Note also that github doesn't show columns correctly, so if you want to inspect the data, we recommend that you download it and open it in a spreadsheet.

As the mistake your noticed (thank you again for being so thorough!) was a typo in the text, and not a mistake in the data, it does not affect any other calculations reported in this paper.

> Figure 1 D and E: The average number of spikes fired as a function of conductance duration and all at 3
different peak conductances. Are there any trends at the single cell level? For example, if a given
control neuron fires relatively few spikes in response to 500 ms duration, does it also tend to fire
relatively fewer spikes with 1000ms duration? Or maybe if a given neuron fires a relatively high number
of spikes with 500ms then it will typically fire less at 1000ms duration? Although the individual
responses are shown in 1D, it is impossible to follow the number of spikes fired by any individual neuron
in response to the different conductance durations.

At least for high-spiking neurons, the answer to this question can actually be inferred from Figure 1D: those curves that are placed high on the graph for duration of 500 ms, also tend to remain high for 1000 ms, indicating that high-spiking neurons tend to be high-spiking across the range of durations. Numerically, the correlation coefficient between the average number of spikes in response to 500 and 1000 ms, across all neurons, is r=0.94. Arguably, this was to be expected as intrinsic excitability of neurons (whether the neuron is spiky or not) is a much stronger feature than the temporal tuning. (Temporal tuning in the tectum was not even described until this paper, while the variation in intrinsic excitability in the tectum is well known for more than 20 years).

@@

> Figure 4 D, E and F: Is “spikiness” synonymous with “N Spikes”? If these terms do not mean the same
thing, they should be defined. If they do mean the same thing, then one term should be chosen and
used consistently throughout.

As described above, "spikiness" is now consistently used across the manuscript as the name of one of three variables used to quantify intrinsic tuning of tectal neurons.

### 2019-05-31 reviews and responses.md

> Page 10 line 8: The “correction for position” is difficult to understand here and in Methods. Why was a
linear regression used – and how is it determined to generate an accurate “correction”? How do the
data look without this correction?

The linear regression is used because it is the simplest way to compensate for a confounding variable (see ANCOVA, or linear models with confounding variables in general). Linear models are the "default" approach in statistics, and are typically used when the exact relation between two variables is unknown, or when the amount of noise is so high that practically there is no difference between linear and conceivable curvilinear regressions. Both situations apply here.

There is no way for us to determine whether our adjustment is "accurate", as there is no experimental way to make more caudal (and thus younger) cells in the tectum older. This situation is not unique however, as it arises in any study that uses linear models, in any field, from ecology to social sciences. When running a study in a human population, if a variable of interest depends on age, one would typically adjust for age by including age as a covariate in the linear model. Here we perform the same standard procedure. After rereading the manuscript, we realized that we used terms "adjustment", "correction", and, in one case, "compensation" interchangeably, which might have led to some confusion. To stress that we are talking about a basic statistical technique of covarate elimination, we now consistently use the term "adjustment" throughout the manuscript.

In the paper, we do not report how the data would have looked without this adjustment for covariates, as it would not be statistically sound to run a linear model with a known covariate without including it. In most cases, it may be expected to underestimate effects for analyses across groups (as the effect of the covarate would be included in the error), and overestimate effects for explanatory linear models in the second half of the paper (as the covariate would contribute to both the dependent, and the independent variables in the model). For each particular comparison, however, both d, F and p values would differ by a small random value. For example, when the measure of "spikness" (one of the three tuning parameters) is compared across groups, a correct analysis with adjustment for position (`smean~ rostral+medial+group`) produces a p-value of 0.01. Without adjustment (smean \~ group) we get a p-value of 0.03. For temporal tuning, correct analysis (sbend \~ rostral+medial+group) gives p=0.0304, and if covariate is left unadjusted (sbend~gropu) we get p=0.029. Similarly, for amplitude tuning, we get p=0.00209 for correct analysis, and p=0.0024 for unadjusted analysis. For the analyses in which temporal tuning and ampliltude coefficients were explained through intrinsic parameters, the optimal models were not qualitatively different between adjusted and unadjusted data, but all values in the table were slightly different.

### 2019-05-31 reviews and responses.md

> Page 10 line 8: The “correction for position” is difficult to understand here and in Methods. Why was a
linear regression used – and how is it determined to generate an accurate “correction”? How do the
data look without this correction?

The linear regression is used because it is the simplest way to compensate for a confounding variable (see ANCOVA, or linear models with confounding variables in general). Linear models are the "default" approach in statistics, and are typically used when the exact relation between two variables is unknown, or when the amount of noise is so high that practically there is no difference between linear and conceivable curvilinear regressions. Both situations apply here.

There is no way for us to determine whether our adjustment is "accurate", as there is no experimental way to make more caudal (and thus younger) cells in the tectum older. This situation is not unique however, as it arises in any study that uses linear models, in any field, from ecology to social sciences. When running a study in a human population, if a variable of interest depends on age, one would typically adjust for age by including age as a covariate in the linear model. Here we perform the same standard procedure. After rereading the manuscript, we realized that we used terms "adjustment", "correction", and, in one case, "compensation" interchangeably, which might have led to some confusion. To stress that we are talking about a basic statistical technique of covarate elimination, we now consistently use the term "adjustment" throughout the manuscript.

In the paper, we do not report how the data would have looked without this adjustment for covariates, as it would not be statistically sound to run a linear model with a known covariate without including it. In most cases, it may be expected to underestimate effects for analyses across groups (as the effect of the covarate would be included in the error), and overestimate effects for explanatory linear models in the second half of the paper (as the covariate would contribute to both the dependent, and the independent variables in the model). For each particular comparison, however, both d, F and p values would differ by a small random value. For example, when the measure of "spikness" (one of the three tuning parameters) is compared across groups, a correct analysis with adjustment for position (`smean~rostral+medial+group`) produces a p-value of 0.01. Without adjustment (`smean~group`) we get a p-value of 0.03. For temporal tuning, correct analysis (`sbend~rostral+medial+group`) gives p=0.0304, and if covariate is left unadjusted (`sbend~group`) we get p=0.029. Similarly, for amplitude tuning, we get p=0.00209 for correct analysis, and p=0.0024 for unadjusted analysis. For the analyses in which temporal tuning and ampliltude coefficients were explained through intrinsic parameters, the optimal models were not qualitatively different between adjusted and unadjusted data, but all values in the table were slightly different.

@@

"Intrinsic spikiness" is no longer used.

"Intrinsic excitability" is used as a name for the phenomenon, that in this particular study we assess mostly using the "spikness" variable, although arguably the number of spikes observed in current clamp experiments can also be considered a different quantification of the same variable. In the text, we say "intrinsic excitability" when we make inferences about the actual properties of the neuron, and "spikiness" when we are talking about the actual value we calculated to assess and quantify this intrinsic excitability.

> Page 15 line 34: The average membrane resistance (330 MOhm) seems a bit low for these small immature
tectal neurons. Were the neurons healthy? What was the average resting membrane potential? Were
there any set criteria for Rm, Vm, or access resistance?

Thank you for catching this error! We had an order of magnitude typo here, and the average membrane resistance was in fact 3.2±1.5 GOhm. This value is a bit higher than reported earlier (1.2±0.8 GOhm in Ciarleglio Khakhalin 2015; about 1.1 in Khakhalin 2011), but within what can be reasonably expected for healthy neurons of this size. To make sure that this number reflects our observations, for the purpose of this response, we verified this value by reanalyzing our data offline, and obtained a similar estimation for the average membrane resistance (2.9±1.4 GOhm). We have several reasons to believe that the cells included in this study were healthy. First, the access resistances obtained were good (30±14 MOhm, which is similar to benchmark 50±23 in Ciarleglio Khakhalin 2015, and about 20-30 MOhm in Khakhalin 2011), indicating that higher membrane resistance is not likely to be due to poor access. Second, our neurons reliably produced spikes in current clamp mode, and active currents in response to membrane depolarization in voltage clamp mode, and the quantification of these active conductances yielded numbers similar to those reported before (Ciarleglio Khakhalin 2015). The quality of voltage clamp, as judged by the absence of escaping active currents in IV protocols, was good in the majority of neurons (82% of all neurons recorded). Finally, a few cells that had their membrane resistance change drastically during the recording session, were discarded at the data collection stage. We did not measure the resting potential directly, but measured the holding current necessary to bring the cell to -60 mV in voltage clamp mode, and the mean value of this I_hold value was -8.2±4.8 pA, which is not higher tha what was previously recorded in the literature (19.8±16.6 in Ciearleglio Khakhalin 2015).
